Urban and socio-economic correlates of property price evolution:~Application to Dublin's costal area

### 2020-08-06 dsaa_application_manuscript/dsaa_application_manuscript.Rmd
renv

# libraries --------------------------------------------------------------------
library(tidyverse)
library(magrittr)
library(here)
library(lubridate)
library(papaja)
library(patchwork)
library(kableExtra)
library(latex2exp)
library(yardstick)
library(sf)
library(tmap)
library(lwgeom)
library(osmdata)
library(OpenStreetMap)
library(mgcv)
library(xgboost)

@@

xgmodel <- xgboost::xgboost(
  data = train_tbl_x,
  label = train_tbl_y,
  objective = "reg:linear",
  eval_metric = "rmse",
  verbose = 0,
  nround = 100
)

@@

lm_result <- lm(pred ~ price, result) %>%
  papaja::apa_print.lm()
```

>>>>

# libraries --------------------------------------------------------------------
library(rticles)
library(tidyverse)
library(magrittr)
library(here)
library(lubridate)
library(papaja)
library(patchwork)
library(kableExtra)
library(latex2exp)
library(yardstick)
library(sf)
library(tmap)
library(lwgeom)
library(osmdata)
library(OpenStreetMap)
library(mgcv)
library(xgboost)

@@

xgmodel <- xgboost::xgboost(
  data = train_tbl_x,
  label = train_tbl_y,
  objective = "reg:squarederror",
  eval_metric = "rmse",
  verbose = 0,
  nround = 100
)

@@

lm_result <- lm(pred ~ price, result) %>%
  papaja::apa_print()
```

### 2020-08-06 dsaa_application_manuscript/dsaa_application_manuscript.tex
renv

@@

Using the XGBoost algorithm, the 160 urban features are explaining 47\% of the property price variance (\(F(1, 2076) = 1,840.91\), \(p < .001\)) with a \(RMSE\) of €124,364 and a \(MAE\) of €83,963.61.

@@

\caption{\label{tab:osm-features-table}Urban features importance (higher than 1\%).}
\centering
\fontsize{8}{10}\selectfont
\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\toprule
Feature Category & Feature Type & Importance\\
\midrule
amenity & embassy & 18.2\%\\
natural & grassland & 6.7\%\\
power & line & 2.3\%\\
route & bus & 2.2\%\\
boundary & administrative & 2.1\%\\
boundary & political & 1.7\%\\
route & ferry & 1.7\%\\
highway & residential & 1.3\%\\
highway & secondary & 1.2\%\\
historic & yes & 1.2\%\\
route & power & 1.1\%\\
cycleway & lane & 1.1\%\\
service & crossover & 1.1\%\\
barrier & wall & 1.1\%\\
place & locality & 1.1\%\\
area & yes & 1.0\%\\
\bottomrule
\end{tabu}
\end{table}

By analysing their importance (Table \ref{tab:osm-features-table}), the most relevant geographical features to predict housing prices are the distance to an embassy (18.2\%) and the distance to natural grasslands such as parks and gardens (6.7\%).

@@

Using the XGBoost algorithm, the 48 socio-economic features explain 42\% of the property price variance (\(F(1, 2076) = 1,508.10\), \(p < .001\)) with a \(RMSE\) of €131,322 and a \(MAE\) of €87,328.79.

@@

\caption{\label{tab:census-features-table}Socio-economic features importance (higher than 1\%).}
\centering
\fontsize{8}{10}\selectfont
\begin{tabular}[t]{lll}
\toprule
Feature Category & Feature Type & Importance\\
\midrule
housing rooms & 8 or more Rooms & 29.0\%\\
religion & No Religion & 5.0\%\\
population & Age 0 - 14 & 3.8\%\\
housing rooms & 6 Rooms & 3.3\%\\
general health & Very Good & 3.1\%\\
population & Age 80 Plus & 2.8\%\\
housing tenure & Owner Occupier with Mortgage & 2.5\%\\
general health & Good & 2.4\%\\
religion & Other Catholic & 2.4\%\\
housing rooms & 5 Rooms & 2.3\%\\
disabilty age group & Persons with a disability aged 25 - 44 & 2.3\%\\
religion & No Answer & 2.2\%\\
housing rooms & 7 Rooms & 2.2\%\\
disabilty age group & Persons with a disability aged 0 - 14 & 2.0\%\\
population & Age 45 - 64 & 1.8\%\\
population & Age 15 - 24 & 1.6\%\\
disabilty age group & Persons with a disability aged 65 Plus & 1.6\%\\
housing rooms & 3 Rooms & 1.6\%\\
housing tenure & Social Rented & 1.6\%\\
disabilty age group & Persons with a disability aged 45 - 64 & 1.5\%\\
housing tenure & Rented Free of Rent & 1.4\%\\
disabilty age group & Persons with a disability aged 15 - 44 & 1.4\%\\
general health & Bad & 1.4\%\\
housing rooms & 1 Room & 1.4\%\\
housing tenure & Owner Occupier No Mortgage & 1.4\%\\
disabilty age group & Total persons with a disability & 1.2\%\\
housing rooms & 4 Rooms & 1.2\%\\
population & Female & 1.2\%\\
religion & Roman Catholic & 1.2\%\\
population & Age 25 - 44 & 1.1\%\\
carers & Provides No Care & 1.1\%\\
housing tenure & Private Rented & 1.0\%\\
\bottomrule
\end{tabular}
\end{table}

According to the analysis of the relative importance of socio-economic features (Table \ref{tab:census-features-table}), the most important socio-economic features are the proportion of large houses (i.e., houses with eight or more rooms) in the small area containing the property (29.0\%). It also appears that areas having a high proportion of young children (3.8\%), as well as the proportion of people reporting that they have no religion (5.0\%), influence the model.

@@

The authors would like to thank the developers of the following R packages used to process, analyse, display and report data: R (Version 4.0.0; R Core Team 2020) and the R-packages \emph{dplyr} (Version 1.0.0; Wickham, François, et al. 2020), \emph{forcats} (Version 0.5.0; Wickham 2020), \emph{ggplot2} (Version 3.3.1; Wickham 2016), \emph{here} (Version 0.1; Müller 2017), \emph{kableExtra} (Version 1.1.0; Zhu 2019), \emph{latex2exp} (Version 0.4.0; Meschiari 2015), \emph{lubridate} (Version 1.7.8; Grolemund and Wickham 2011), \emph{lwgeom} (Version 0.2.4; Pebesma 2020), \emph{magrittr} (Version 1.5; Bache and Wickham 2014), \emph{mgcv} (Version 1.8.31; Wood 2011, 2004, 2003; Wood et al. 2016), \emph{nlme} (Version 3.1.147; Pinheiro et al. 2020), \emph{OpenStreetMap} (Version 0.3.4; Fellows and JMapViewer library by Jan Peter Stotz 2019), \emph{osmdata} (Version 0.1.3; Padgham et al. 2017), \emph{papaja} (Version 0.1.0.9942; Aust and Barth 2020), \emph{patchwork} (Version 1.0.0; Pedersen 2019), \emph{purrr} (Version 0.3.4; Henry and Wickham 2020), \emph{readr} (Version 1.3.1; Wickham, Hester, and Francois 2018), \emph{sf} (Version 0.9.3; Pebesma 2018), \emph{stringr} (Version 1.4.0; Wickham 2019), \emph{tibble} (Version 3.0.1; Müller and Wickham 2020), \emph{tidyr} (Version 1.1.0; Wickham and Henry 2020), \emph{tidyverse} (Version 1.3.0; Wickham, Averick, et al. 2019), \emph{tmap} (Version 3.0; Tennekes 2018), \emph{xgboost} (Version 1.0.0.2; Chen et al. 2020), and \emph{yardstick} (Version 0.0.6; Kuhn and Vaughan 2020).

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-antonakakis2016dynamic}{}%
Antonakakis, Nikolaos, and Christos Floros. 2016. ``Dynamic Interdependencies Among the Housing Market, Stock Market, Policy Uncertainty and the Macroeconomy in the United Kingdom.'' \emph{International Review of Financial Analysis} 44: 111--22.

@@

\leavevmode\hypertarget{ref-R-kableExtra}{}%
Zhu, Hao. 2019. \emph{KableExtra: Construct Complex Table with 'Kable' and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.

>>>>

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{cslreferences}%
  {\setlength{\parindent}{0pt}%
  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
  {\par}

@@

Using the XGBoost algorithm, the 160 urban features are explaining 44\% of the property price variance (\(F(1, 2076) = 1,629.56\), \(p < .001\)) with a \(RMSE\) of €127,783 and a \(MAE\) of €85,115.25.

@@

\caption{\label{tab:osm-features-table}Urban features importance (higher than 1\%).}
\centering
\fontsize{8}{10}\selectfont
\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\toprule
Feature Category & Feature Type & Importance\\
\midrule
amenity & embassy & 17.7\%\\
natural & grassland & 6.7\%\\
route & bus & 2.3\%\\
power & line & 2.2\%\\
boundary & administrative & 1.5\%\\
place & island & 1.5\%\\
barrier & wall & 1.5\%\\
cycleway & share & 1.4\%\\
route & ferry & 1.3\%\\
cutting & yes & 1.1\%\\
amenity & bar & 1.1\%\\
barrier & yes & 1.1\%\\
place & locality & 1.1\%\\
boundary & political & 1.1\%\\
route & road & 1.1\%\\
highway & path & 1.1\%\\
tunnel & yes & 1.0\%\\
boundary & postal & 1.0\%\\
\bottomrule
\end{tabu}
\end{table}

By analysing their importance (Table \ref{tab:osm-features-table}), the most relevant geographical features to predict housing prices are the distance to an embassy (17.7\%) and the distance to natural grasslands such as parks and gardens (6.7\%).

@@

Using the XGBoost algorithm, the 48 socio-economic features explain 42\% of the property price variance (\(F(1, 2076) = 1,500.87\), \(p < .001\)) with a \(RMSE\) of €131,270 and a \(MAE\) of €86,891.85.

@@

\caption{\label{tab:census-features-table}Socio-economic features importance (higher than 1\%).}
\centering
\fontsize{8}{10}\selectfont
\begin{tabular}[t]{lll}
\toprule
Feature Category & Feature Type & Importance\\
\midrule
housing rooms & 8 or more Rooms & 28.2\%\\
population & Age 0 - 14 & 4.5\%\\
religion & No Religion & 4.4\%\\
population & Age 80 Plus & 3.6\%\\
general health & Very Good & 3.5\%\\
housing rooms & 6 Rooms & 3.2\%\\
housing tenure & Owner Occupier with Mortgage & 2.5\%\\
disabilty age group & Persons with a disability aged 65 Plus & 2.3\%\\
housing rooms & 5 Rooms & 2.2\%\\
housing tenure & Social Rented & 2.2\%\\
housing rooms & 4 Rooms & 2.1\%\\
religion & Other Catholic & 2.0\%\\
housing rooms & 7 Rooms & 1.9\%\\
housing rooms & 3 Rooms & 1.8\%\\
religion & No Answer & 1.8\%\\
housing tenure & Private Rented & 1.6\%\\
population & Age 45 - 64 & 1.5\%\\
population & Age 25 - 44 & 1.5\%\\
housing tenure & Owner Occupier No Mortgage & 1.5\%\\
disabilty age group & Persons with a disability aged 0 - 14 & 1.5\%\\
disabilty age group & Persons with a disability aged 25 - 44 & 1.4\%\\
housing rooms & 1 Room & 1.4\%\\
carers & Provides No Care & 1.3\%\\
religion & Roman Catholic & 1.3\%\\
general health & Good & 1.3\%\\
population & Female & 1.3\%\\
general health & Fair & 1.3\%\\
disabilty age group & Total persons with a disability & 1.2\%\\
general health & Bad & 1.2\%\\
religion & Other Religion & 1.2\%\\
disabilty age group & Persons with a disability aged 45 - 64 & 1.2\%\\
carers & Total Care Providers & 1.2\%\\
population & Age 15 - 24 & 1.1\%\\
carers & 20-49 hours unpaid PW & 1.1\%\\
housing tenure & Rented Free of Rent & 1.0\%\\
\bottomrule
\end{tabular}
\end{table}

According to the analysis of the relative importance of socio-economic features (Table \ref{tab:census-features-table}), the most important socio-economic features are the proportion of large houses (i.e., houses with eight or more rooms) in the small area containing the property (28.2\%). It also appears that areas having a high proportion of young children (4.4\%), as well as the proportion of people reporting that they have no religion (4.5\%), influence the model.

@@

The authors would like to thank the developers of the following R packages used to process, analyse, display and report data: R (Version 4.0.2; R Core Team 2020) and the R-packages \emph{dplyr} (Version 1.0.1; Wickham, François, et al. 2020), \emph{forcats} (Version 0.5.0; Wickham 2020), \emph{ggplot2} (Version 3.3.2; Wickham 2016), \emph{here} (Version 0.1; Müller 2017), \emph{kableExtra} (Version 1.1.0; Zhu 2019), \emph{latex2exp} (Version 0.4.0; Meschiari 2015), \emph{lubridate} (Version 1.7.9; Grolemund and Wickham 2011), \emph{lwgeom} (Version 0.2.5; Pebesma 2020), \emph{magrittr} (Version 1.5; Bache and Wickham 2014), \emph{mgcv} (Version 1.8.31; Wood 2011, 2004, 2003; Wood et al. 2016), \emph{nlme} (Version 3.1.148; Pinheiro et al. 2020), \emph{OpenStreetMap} (Version 0.3.4; Fellows and JMapViewer library by Jan Peter Stotz 2019), \emph{osmdata} (Version 0.1.3; Padgham et al. 2017), \emph{papaja} (Version 0.1.0.9997; Aust and Barth 2020), \emph{patchwork} (Version 1.0.1; Pedersen 2019), \emph{purrr} (Version 0.3.4; Henry and Wickham 2020), \emph{readr} (Version 1.3.1; Wickham, Hester, and Francois 2018), \emph{sf} (Version 0.9.5; Pebesma 2018), \emph{stringr} (Version 1.4.0; Wickham 2019), \emph{tibble} (Version 3.0.3; Müller and Wickham 2020), \emph{tidyr} (Version 1.1.0; Wickham and Henry 2020), \emph{tidyverse} (Version 1.3.0; Wickham, Averick, et al. 2019), \emph{tmap} (Version 3.1; Tennekes 2018), \emph{xgboost} (Version 1.1.1.1; Chen et al. 2020), and \emph{yardstick} (Version 0.0.7; Kuhn and Vaughan 2020).

@@

\hypertarget{refs}{}
\begin{cslreferences}
\leavevmode\hypertarget{ref-antonakakis2016dynamic}{}%
Antonakakis, Nikolaos, and Christos Floros. 2016. ``Dynamic Interdependencies Among the Housing Market, Stock Market, Policy Uncertainty and the Macroeconomy in the United Kingdom.'' \emph{International Review of Financial Analysis} 44: 111--22.

@@

\leavevmode\hypertarget{ref-R-kableExtra}{}%
Zhu, Hao. 2019. \emph{KableExtra: Construct Complex Table with 'Kable' and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.
\end{cslreferences}
